Capon Beamforming Applied on Moving Objects - An Analysis of Local Lateral Shift-Invariance

### 2013-12-18 HRUIofMovingObjects/jrnl_article.tex
Moving objects. Almost there-....

\IEEEPARstart{I}{n} the last decade several authors have applied the Capon beamformer (also known as the minimum variance beamformer) for medical ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The method has been shown to improve the lateral resolution and edge sharpness \cite{Synnevag2007, Synnevag2009, Chen2011}. This is accomplished by adaptively adjusting the array weights based on element-to-element covariance in order to minimize interference. The method is, however, also highly sensitive to model errors \cite{Mestre2006, Widrow1982, Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying steering vector uncertainty sets \cite{Lorenz2005, Rubsamen2013}. Spatial smoothing and diagonal loading, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution parameters \cite{Synnevag2007a}. Steering vector uncertainty sets, however, have not been applied for medical ultrasound imaging.

@@

A steering vector uncertainty set reduce signal-nulling which appears if the signal of interest is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the Capon beamformer obtains its super resolution capabilities, but the signal can also get lost in the process. In \cite{JianLi2003} it was shown that most methods based on steering vector uncertainty sets are equivalent to diagonal loading where the loading factor depends on the steering vector uncertainty. In that way they make up for a given uncertainty by making the Capon beam wider. Hence, the maximal achievable resolution is decreased if the initial steering vector spacing was too coarse. In medical ultrasound imaging this initial set of steering vectors is governed by the number of receive beams. In this paper we will make use of an additional set of steering vectors to reduce the self-nulling effect by means of oversampling. The maximal achievable resolution is then maintained without the presence of severe self-nulling.

In \cite{Cox1973}, Cox presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. He also states the following important fact about the Capon beamformer [the $\vec{k}_3$ processor]: 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to four times the array resolution using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the high precision scanned output when the object is located at plus and minus half the required beam spacing for DAS [See (\ref{eq:resolution})]. %No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. 
The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) has been added to the Capon beamformer output. A diagonal loading of $1/100$ corresponds to a signal-to-noise ratio (SNR) of 20 dB before array gain is applied. The figure shows that a DAS beamformer, using the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However if the Capon beamformer is used with the same beam density the loss could be as large as $27$ dB. This imaging system is then highly sensitive to small spatial shifts, either by the object or the array itself. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam spacing or object distance is decreased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Theoretical drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming ($M=64$, $L=32$, $d=0.01$) applied in a farfield-narrowband setting. %The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the required beam spacing for delay-and-sum, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.
}
\label{fig:das_capon_beams}
\end{figure}

@@

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). The dotted vertical lines indicates the position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} with a probe setup equal to the M5S probe by GE Vingmed Ultrasound, Horten, Norway. The probe was excited with a 2.5 MHz and 1.5 cycles long pulse. The transmit focus in azimuth and elevation were configured to be equal and the point scatterer was moved through this point. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. %In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

@@

In Fig.\,\ref{fig:das_shiftvariant} the beam density is reduced by a factor of two compared to Fig.\,\ref{fig:das_shiftinvariant}. We see how the geometric distortion increases to a level where the contours no longer are diagonal. The maximum scalloping loss is also larger than 1 dB. Hence, the imaging system is no longer locally laterally shift-invariant. We observe that the maximum peak is both geometrically distorted and attenuated when the point is located between two beams and the distance between the beams are too high.

@@

\section{Local Lateral Shift-Invariance of the Capon Beamformer}\label{sec:capon_LLSI}
We will now investigate the local lateral shift-invariance of the Capon beamformer with parameters equal to the ones used in earlier work \cite{Synnevag2009, Asen}. In Fig.\,\ref{fig:capon1} the same point scatterer response as presented in Fig.\,\ref{fig:das_shiftinvariant} has been weighted with Capon weights calculated using (\ref{eq:capon_weights}) with $L = M/2 = 48$, $K=1$ and $d=1/100$. The lateral offsets are calculated in such a way that the beam at zero degrees is the only one where the point scatterer is located in the center of the beam. The signal cancellation caused by the difference between the signal propagation vector and the assumed steering vector is therefore minimal when the point scatterer is located at this position. From this plot we observe that the signal cancellation is high for all the other offsets. As a consequence the scalloping loss is tremendous, and not far from the predicted value in Fig\,\ref{fig:das_capon_beams}. %The difference between the predicted and the simulated loss might be caused by several factors. The most important one being that Fig.\,\ref{fig:das_capon_beams} shows a single frequency whereas the RMS operator used in Fig.\,\ref{fig:capon1} average a broadband signal in range. For a slice through the center of the pulse the scalloping loss is below -30dB.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig3a.eps}\label{fig:capon1}
}
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig3b.eps}\label{fig:capon2}
}
}
\caption{LSV-plots of Capon beamforming with $L=M/2=48$ and $K=1$, and with sampling corresponding to (\ref{eq:resolution}). a) Capon beamforming with $d=1/100$ as diagonal loading. The system is not locally lateral shift-invariant. b) The diagonal loading factor $d$ is increased to 1. The system is still not locally lateral shift-invariant.}
\label{fig:capon}
\end{figure*}

\section{Oversampling Methods}\label{sec:methods}
In this section we will discuss different methods aimed at improving the local lateral shift-invariance for the Capon beamformer. From the previous section it should be clear that we somehow need to increase the lateral sampling to make the Capon imaging system locally laterally shift-invariant. In this section we will investigate how this oversampling should be conducted, how large it needs to be and how the proposed methods will affect the system frame rate and the processing complexity.

@@

From Fig.\,\ref{fig:capon_oversampling_b} we see that sixteen times oversampling is not enough to achieve a local lateral shift-invariant Capon beamformer when a moving point scatterer is imaged without noise and with the same parameters as in Fig\,\ref{fig:capon}, but we are getting closer. %$1/100$ in diagonal loading. 
Note again that a diagonal loading factor of $1/100$ will introduce a white noise component in the sample covariance matrix $20$ dB lower than the signal present. Hence, the covariance has been estimated from data which appeared to have an SNR of 20 dB.

Even though increasing the number of transmits improves the lateral shift-invariance it will also reduce the imaging frame rate with a factor equal to the oversampling factor. The processing complexity will also increase equally. For these two reasons, oversampling on transmit is something which is desirable to avoid in real time ultrasound applications, and therefore we will not discuss this method further.

@@

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams high while the number of tx-beams are reduced. When using PRB for oversampling we would like to maintain the same number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand more rx-beams will increase the number of covariance matrices, and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Since more than ten times oversampling is required in order to achieve lateral shift invariant high resolution imaging of a moving point scatterer \textit{in-vitro}, none of the implementations described in \cite{Asen} would be real time any more. We therefore need a different way of oversampling our imaging sector.

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since ultrasound data have to be pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole imaging sector by phase steering. This is also possible in broadband applications as long as the phase rotation is less than one pulse length (See coarse-fine beamforming in \cite{Thomenius}). The maximal steering angle is therefore given by
\begin{align}
\theta_{max} = \alpha p \Delta,
\end{align}
where $\alpha$ is the pulse length in $\lambda$'s, and $p$ and $\Delta$ is the oversampling factor and beam spacing respectively as defined in (\ref{eq:resolution}). For our simulations $\theta_{max} = 3\Delta$.
It is therefore possible to make a sweep using phase rotation between a given beam and half-way to its two nearest neighbours without introducing large errors. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

@@

Since the phase rotation method operates on the same channel data as in Fig.\,\ref{fig:capon1}, there will be no reduction in the acquisition frame rate by applying this method. In addition, since the same number of covariance matrices are constructed and inverted, there is no extra cost involved here as well. The only additional cost will be the multiplication of the extra steering vectors with the inverted covariance matrix. If the equation $\mat{\hat{R}}\vec{a} = \vec{b}$, as in \cite{Asen}, is solved using the Gauss-Jordan algorithm, we can solve the problem for all steering vectors at once by row reducing the matrix $[\mat{\hat{R}},  \vec{a}_{\theta_1}, \vec{a}_{\theta_2}, \cdots \vec{a}_{\theta_{N_\theta}}]$ to the following form $[\mat{I},  \vec{b}_{\theta_1}, \vec{b}_{\theta_2}, \cdots \vec{b}_{\theta_{N_\theta}}]$. This is still an $O(L^3)$ operation as long as $N_\theta$ is not much greater than $L$.

@@

Figure \ref{fig:phantom_capon_LSV} shows the same point scatterer response processed with the Capon beamformer without oversampling. For the frame where the point scatterer is at -6 degrees the receive beam direction is very close to the signal propagation vector. Hence, we get a high amplitude value at this location. In almost all the other frames the signal loss is massive (10 dB). The result is blinking point scatterers when subjected to motion \multimedia{Video Capon}.

Figure \ref{fig:phantom_capon_LSV_10PR} presents the result of applying sixteen times oversampling by phase rotation. We observe how the amplitude of the point scatterer is better preserved in all frames, and the scalloping loss is not far from our 1 dB threshold. Note that the oscillation along the diagonal does not indicate shift-variance. It is a result of lack of control when the probe was manually dragged across the phantom. In order to get a diagonal without peaks the probe has to be dragged across the phantom with a speed which matches the acquisition frame rate. In addition, due to the increase in lateral resolution, the Capon beamformer requires smaller frame-to-frame movements than DAS. Looking at the video of Capon beamforming with ten times oversampling by phase rotation \multimedia{Video Capon + 10PR} the blinking is now highly reduced. %The system is however not fully shift invariant. To achieve this a higher level of oversampling, %We therefore conclude that the system is now locally laterally shift invariant.

@@

\section{Discussion}\label{sec:dis}
In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. We have shown that the Capon beamformer is less sensitive to small lateral shifts if oversampling based on phase rotation is applied. This without reducing the acquisition frame rate and with a minor increase in processing complexity. %In Section \ref{sec:capon_LLSI} it was shown that a Capon beamformer with high resolution parameters is not locally laterally shift-invariant when subjected to motion. The root cause was found to be too coarse lateral sampling. 
%In Section \ref{sec:methods} we presented methods for improved shift-invariance based on oversampling. The most promising method, with respect to improved shift-invariance, reduction in acquisition frame rate and processing complexity combined, was based on phase rotation. 
%In Section \ref{sec:res} we presented results showing that this method also works on \textit{in-vitro} data when the ultrasound probe is subjected to motion. 
In simulations the method reduced the maximal scalloping loss by more than 20 dB. On \textit{in-vitro} data the reduction was more than 7 dB. On \textit{in-vitro} data the imaging system is almost locally laterally shift invariant.

Note that oversampling is not only essential to keep the amplitude level of moving point scatterers. It should also helps to improve contrast and edge definitions. If oversampling is not applied homogeneous speckle regions could get a too low average value and the edge delineation of a moving object will fluctuate in the same way as a moving point scatterer would have sparkled.

@@

Faced with an imaging scene containing bright point scatterers or sharp edges, one could wonder how large the oversampling factor needs to be in order to avoid visible scalloping losses. In his seminal paper, Cox \cite{Cox1973} derives the required beam density for a given output SNR. For instance 47 dB requires a beam density for the Capon beamformer ten times the one given by (\ref{eq:resolution}). Since resolution for the Capon beamformer depends on SNR we would recommend to first use the dynamic range in the image as guidance. The amount of diagonal loading will also dictate the maximal achievable resolution since $10\log{d}$ can be interpreted as the maximal input SNR seen by the Capon beamformer. %Cox measures resolution as the 3 dB saddle point between two impinging monochromatic plane waves. For ultrasound data only a 50 \% increase in resolution was found in \cite[Fig.~7]{Asen} for a 30 dB output SNR when measured in a similar way as Cox [6 dB saddle point]. However, sixteen times oversampling was required to make the imaging system shift invariant in the same paper. In this paper more than ten times oversampling was found to be required for a 40 dB output SNR (after array gain) in simulations, however for the \textit{in-vitro} data ten times oversampling was enough for the system to be locally laterally shift-invariant. This might be a result of improved initial sampling (as discussed in Section \ref{sec:res}) and a higher initial noise level before diagonal loading was applied.

Relating the presented phase rotation method to earlier work we find that it is similar to a steering vector uncertainty set. However, where these set-based methods typically has a criterion for selecting the best output after applying all the steering vectors contained in the set, oversampling with phase rotation outputs all. It can also be thought of as a narrow band approach to parallel receive beamforming where Capon is applied on all receive beams.

Oversampling by phase rotation can therefore introduce similar block artifacts as parallel receive beamforming (PRB). These artifacts should be possible to remove in the same way as for PRB by using the synthetic transmit beamforming (STB) technique. If we distribute our steering angles in an overlapping scheme [like rx-beams are spread out for STB] we get a phase rotation version of STB where e.g. eight phase rotated beams are averaged into four. The geometric distortion, or stripes, seen at the bottom of the sector in the video \multimedia{Capon + 10PR} should be possible to remove using this technique.

When oversampling is applied it becomes evident that the high resolution settings used in this paper might be to extreme, both with respect to the level of required oversampling and the point's appearance. Traditionally in ultrasound imaging the point spread function is always wider laterally than in range. With the Capon beamformer this is now changed. A more symmetric point spread function while preserving speckle statistics is obtained with the parameters $L=M/4$ and $K=0$ \cite{Synnevag2007a}. The level of required oversampling and the computational complexity are also reduced. %Another option is to perform pulse inversion if a spiky symmetric point spread function is desirable.

Finally we want to stress the fact that this demand for oversampling does not only affect the Capon [or minimum variance beamformer]. Other methods claiming super resolution, like the Eigenspace method, low complexity beamformer etc. \cite{Synnevag2011, Nilsen2009, Mehdizadeh2012a, Kim}, would also need oversampling when higher lateral resolution is obtained. The same applies for high resolution spectral estimation \cite{Ekroll2012}.

@@

\section{Conclusion}\label{sec:con}
In this paper we have investigated the lateral shift-invariance properties of the Capon beamformer when imaging moving objects. It was shown that Capon beamforming with the same beam density as required for delay-and-sum beforming results in blinking point scatterers when subjected to motion.  We then discussed and investigated different methods in order to reduced this effect. Finally we proved that the local lateral shift-invariance could be improved by oversampling based on phase rotation without affecting the acquisition frame rate and with minor changes in processing complexity.

>>>>

\IEEEPARstart{I}{n} the last decade several authors have applied the Capon beamformer (also known as the minimum variance beamformer) for medical ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The method has been shown to improve lateral resolution and edge sharpness \cite{Synnevag2007, Synnevag2009, Chen2011} in ultrasound images. This is accomplished by adaptively adjusting the array weights based on element-to-element covariance in order to minimize interference. The method is, however, also highly sensitive to model errors \cite{Mestre2006, Widrow1982, Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying steering vector uncertainty sets \cite{Lorenz2005, Rubsamen2013}. Spatial smoothing and diagonal loading, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution parameters \cite{Synnevag2007a}. Steering vector uncertainty sets, however, have not been applied for medical ultrasound imaging.

@@

A steering vector uncertainty set reduces signal-nulling which appears if the signal of interest is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the Capon beamformer obtains its super resolution capabilities, but the signal can also get lost in the process. In \cite{JianLi2003} it was shown that most methods based on steering vector uncertainty sets are equivalent to diagonal loading where the loading factor depends on the steering vector uncertainty. In that way they make up for a given uncertainty by making the Capon beam wider. Hence, the maximal achievable resolution is decreased if the initial steering vector spacing was too coarse. In medical ultrasound imaging this initial set of steering vectors is governed by the number of receive beams. In this paper we will make use of an additional set of steering vectors to reduce the self-nulling effect by means of oversampling. The maximal achievable resolution is then maintained without the presence of severe self-nulling.

In \cite{Cox1973}, Cox presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. He also states the following important fact about the Capon beamformer [the $\vec{k}_3$ processor]: 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to four times the array resolution using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the high precision scanned output when the object is located at plus and minus half the required beam spacing for DAS [See (\ref{eq:resolution})]. %No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. 
The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) have been added to the Capon beamformer output. We see that a DAS beamformer, using the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However if the Capon beamformer is used with the same beam density the loss could be as large as $27$ dB. This imaging system is then highly sensitive to small spatial shifts, either by the object or the array itself. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam spacing or object distance is decreased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Estimated drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming ($M=64$, $L=32$, $d=0.01$) applied in a farfield-narrowband setting. %The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the required beam spacing for delay-and-sum, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.
}
\label{fig:das_capon_beams}
\end{figure}

@@

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). Dotted vertical lines indicate position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} with a probe setup equal to the M5S probe by GE Vingmed Ultrasound, Horten, Norway. The probe was excited with a 2.5 MHz and 1.5 cycles long pulse. The transmit focus in azimuth and elevation were configured to be equal and the point scatterer was moved through this point. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. %In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

@@

In Fig.\,\ref{fig:das_shiftvariant} the beam density is reduced by a factor of two compared to Fig.\,\ref{fig:das_shiftinvariant}. The geometric distortion is now increased to a level where the contours are not diagonal. The maximum scalloping loss is also larger than 1 dB. Hence, the imaging system is no longer locally laterally shift-invariant. We observe that the maximum peak is both geometrically distorted and attenuated when the point is located between two beams and the distance between the beams is too high.

@@

\section{Local Lateral Shift-Invariance of the Capon Beamformer}\label{sec:capon_LLSI}
We will now investigate the local lateral shift-invariance of the Capon beamformer with high resolution parameters \cite{Synnevag2009, Asen}. In Fig.\,\ref{fig:capon1} the same point scatterer response as presented in Fig.\,\ref{fig:das_shiftinvariant} has been weighted with Capon weights calculated using (\ref{eq:capon_weights}) with $L = M/2 = 48$, $K=1$ and $d=1/100$. We see how the signal cancellation caused by the difference between the signal propagation vector and the assumed steering vector increases as the point moves away from a given transmit-receive beam. As a consequence the scalloping loss is tremendous, and not far from the predicted value in Fig\,\ref{fig:das_capon_beams}. %The difference between the predicted and the simulated loss might be caused by several factors. The most important one being that Fig.\,\ref{fig:das_capon_beams} shows a single frequency whereas the RMS operator used in Fig.\,\ref{fig:capon1} average a broadband signal in range. For a slice through the center of the pulse the scalloping loss is below -30dB.

@@

\begin{figure*}[!t]
\centerline{
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig3a.eps}\label{fig:capon1}
}
\subfloat[]{
\includegraphics[width=0.5\linewidth]{\img fig3b.eps}\label{fig:capon2}
}
}
\caption{LSV-plots of Capon beamforming with $L=M/2=48$ and $K=1$, and with sampling corresponding to (\ref{eq:resolution}). a) Capon beamforming with $d=1/100$ as diagonal loading. The system is not locally lateral shift-invariant. b) The diagonal loading factor $d$ is increased to 1. The scalloping loss is reduced, but the system is still not locally lateral shift-invariant.}
\label{fig:capon}
\end{figure*}

\section{Oversampling Methods}\label{sec:methods}
In this section we will discuss different methods aimed at improving the local lateral shift-invariance for the Capon beamformer. From the previous section it should be clear that we somehow need to increase the lateral sampling to make the Capon imaging system locally laterally shift-invariant. In this section we will investigate how this oversampling should be conducted, how large it needs to be and how the proposed methods will affect the acquisition frame rate and the processing complexity.

@@

From Fig.\,\ref{fig:capon_oversampling_b} we see that sixteen times oversampling reduces the scalloping loss significantly, but the maximum is still larger than 1 dB. A even higher level of oversampling is therefore required in order to have a completely local lateral shift-invariant Capon beamformer when a moving point scatterer is simulated without noise and with the same parameters as in Fig\,\ref{fig:capon}. %$1/100$ in diagonal loading. 
Note that a diagonal loading factor of $1/100$ will introduce a white noise component in the sample covariance matrix $20$ dB lower than the signal present. Hence, the covariance has been estimated from data which appeared to have an SNR of 20 dB.

Even though increasing the number of transmits improves the lateral shift-invariance it will also reduce the acquisition frame rate with a factor equal to the oversampling factor. The processing complexity will also increase equally. For these two reasons, oversampling on transmit is something which is desirable to avoid in real time ultrasound applications, and therefore we will not discuss this method further.

@@

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams high while the number of tx-beams are reduced. When using PRB for oversampling we would like to maintain the same number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand more rx-beams will increase the number of covariance matrices, and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Since more than sixteen times oversampling is required in order to achieve lateral-shift-invariant and high-resolution imaging of a moving point scatterer, none of the implementations described in \cite{Asen} would be real time any more. We therefore need a different way of oversampling our imaging sector.

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since ultrasound data have to be pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole imaging sector by phase steering. This is also possible in broadband applications as long as the phase rotation is less than one pulse length (See coarse-fine beamforming in \cite{Thomenius}). The maximal steering angle is therefore given by
\begin{align}
\theta_{max} = \alpha p \Delta,
\end{align}
where $\alpha$ is the pulse length in $\lambda$'s, and $p$ and $\Delta$ is the oversampling factor and beam spacing respectively as defined in (\ref{eq:resolution}). For our simulations $\theta_{max} = 3\Delta$.
It is therefore possible to make a sweep using phase rotation between a given beam and half-way to its two nearest neighbours without introducing to large errors. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

@@

Since the phase rotation method operates on the same channel data as in Fig.\,\ref{fig:capon1}, there will be no reduction in the acquisition frame rate by applying this method. In addition, since the same number of covariance matrices are constructed and inverted, there is no extra cost involved here as well. The only additional cost will be the multiplication of the extra steering vectors with the inverted covariance matrix. If the equation $\mat{\hat{R}}\vec{a} = \vec{b}$, as in \cite{Asen}, is solved using the Gauss-Jordan algorithm, we can solve the problem for all steering vectors at once by row reducing the matrix $[\mat{\hat{R}},  \vec{a}_{\theta_1}, \vec{a}_{\theta_2}, \cdots \vec{a}_{\theta_{N_\theta}}]$ to the following form $[\mat{I},  \vec{b}_{\theta_1}, \vec{b}_{\theta_2}, \cdots \vec{b}_{\theta_{N_\theta}}]$. This is an $O(L^3)$ operation as long as $N_\theta$ is not much greater than $L$.

@@

Figure \ref{fig:phantom_capon_LSV} shows the same point scatterer response processed with the Capon beamformer without oversampling. For the frame where the point scatterer is at -6 degrees the receive beam direction is very close to the signal propagation vector. Hence, we see a high amplitude value at this location. In almost all the other frames the signal loss is massive (at most 10 dB). The result is blinking point scatterers when subjected to motion \multimedia{Video Capon}.

Figure \ref{fig:phantom_capon_LSV_10PR} presents the result of applying sixteen times oversampling by phase rotation. We observe how the amplitude of the point scatterer is better preserved in all frames, and the scalloping loss is not far from our 1 dB threshold. Note that the oscillation along the diagonal does not indicate shift-variance. It is a result of lack of control when the probe was manually dragged across the phantom. In order to get a diagonal without peaks the probe has to be dragged across the phantom with a speed which matches the acquisition frame rate. In addition, due to the increase in lateral resolution, the Capon beamformer requires smaller frame-to-frame movements than DAS. Looking at the video of Capon beamforming with sixteen times oversampling by phase rotation \multimedia{Video Capon + 16PR} the blinking is now highly reduced. %The system is however not fully shift invariant. To achieve this a higher level of oversampling, %We therefore conclude that the system is now locally laterally shift invariant.

@@

\section{Discussion}\label{sec:dis}
In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. We have shown that the Capon beamformer is less sensitive to small lateral shifts if oversampling based on phase rotation is applied. This without reducing the acquisition frame rate and with only a minor increase in processing complexity. %In Section \ref{sec:capon_LLSI} it was shown that a Capon beamformer with high resolution parameters is not locally laterally shift-invariant when subjected to motion. The root cause was found to be too coarse lateral sampling. 
%In Section \ref{sec:methods} we presented methods for improved shift-invariance based on oversampling. The most promising method, with respect to improved shift-invariance, reduction in acquisition frame rate and processing complexity combined, was based on phase rotation. 
%In Section \ref{sec:res} we presented results showing that this method also works on \textit{in-vitro} data when the ultrasound probe is subjected to motion. 
In simulations the method reduced the maximal scalloping loss by 16 dB. On \textit{in-vitro} data the reduction was more than 7 dB, and the imaging system is now almost locally laterally shift invariant.

Note that oversampling is not only essential to keep the amplitude level of moving point scatterers. It should also help to improve contrast and edge definitions. If oversampling is not applied homogeneous speckle regions could get a too low average value and the edge delineation of a moving object will fluctuate in the same way as a moving point scatterer would have sparkled.

@@

Faced with an imaging scene containing bright point scatterers or sharp edges, one could wonder how large the oversampling factor needs to be in order to avoid visible scalloping losses. In his seminal paper, Cox \cite{Cox1973} derives the required beam density for a given output SNR. For instance 47 dB requires a beam density for the Capon beamformer ten times the one given by (\ref{eq:resolution}). Since resolution for the Capon beamformer depends on SNR we would recommend to first use the dynamic range in the image as guidance. The amount of diagonal loading will also dictate the maximal achievable resolution since $10\log{d}$ can be interpreted as the maximal input SNR seen by the Capon beamformer. Fig\,\ref{fig:das_capon_beams} shows that 25 times oversampling is required for a $d$-value of $1/100$. When \textit{in-vitro} data contains additional noise this oversampling requirement is reduced.%Cox measures resolution as the 3 dB saddle point between two impinging monochromatic plane waves. For ultrasound data only a 50 \% increase in resolution was found in \cite[Fig.~7]{Asen} for a 30 dB output SNR when measured in a similar way as Cox [6 dB saddle point]. However, sixteen times oversampling was required to make the imaging system shift invariant in the same paper. In this paper more than ten times oversampling was found to be required for a 40 dB output SNR (after array gain) in simulations, however for the \textit{in-vitro} data ten times oversampling was enough for the system to be locally laterally shift-invariant. This might be a result of improved initial sampling (as discussed in Section \ref{sec:res}) and a higher initial noise level before diagonal loading was applied.

Relating the presented phase rotation method to earlier work we find that it is similar to a steering vector uncertainty set. However, where these set-based methods typically has a criterion for selecting the best output after applying all the steering vectors contained in the set, oversampling with phase rotation outputs all. It can also be thought of as a narrow band approach to parallel receive beamforming where Capon's method is applied on all receive beams.

Oversampling by phase rotation can therefore introduce similar block artifacts as parallel receive beamforming (PRB). These artifacts should be possible to remove in the same way as for PRB by using the synthetic transmit beamforming (STB) technique. If we distribute our steering angles in an overlapping scheme [like rx-beams are spread out for STB] we get a phase rotation version of STB where e.g. eight phase rotated beams are averaged into four. The geometric distortion, or stripes, seen at the bottom of the sector in the video with sixteen times oversampling \multimedia{Capon+16PR} should be possible to remove using this technique.

When oversampling is applied it becomes evident that the high resolution settings used in this paper might be to extreme, both with respect to the level of required oversampling and the point scatterer's appearance. Traditionally in medical ultrasound imaging the point spread function is always wider laterally than in range. With the Capon beamformer this is now changed. A more symmetric point spread function while preserving speckle statistics is obtained with the parameters $L=M/4$ and $K=0$ \cite{Synnevag2007a}. The level of required oversampling and the computational complexity are also reduced. %Another option is to perform pulse inversion if a spiky symmetric point spread function is desirable.

Finally we want to stress the fact that this demand for oversampling does not only affect the Capon [or minimum variance beamformer]. Other methods claiming super resolution, like the Eigenspace method, low complexity beamformer etc. \cite{Synnevag2011, Nilsen2009, Mehdizadeh2012a, Kim}, would also require oversampling when improved lateral resolution is obtained. The same applies for high resolution spectral estimation \cite{Ekroll2012}.

@@

\section{Conclusion}\label{sec:con}
In this paper we have investigated the lateral shift-invariance properties of the Capon beamformer when imaging moving objects. It was shown that Capon beamforming with the same beam density as required for delay-and-sum beforming results in blinking point scatterers when subjected to motion.  We then discussed and investigated different methods in order to reduced this effect. Finally we showed that the local lateral shift-invariance could be improved by oversampling based on phase rotation without affecting the acquisition frame rate and with minor changes in processing complexity.
